---
output: 
  word_document:
    reference_docx: ETO_summary_template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```

```{r data, include=FALSE,warning=FALSE}
library(xlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(cowplot)
library(knitr)
library(maps)

Proj_Pop<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/Project_info_for_Summaries.csv",stringsAsFactors = FALSE)
State_adj<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/State_info_for_Summaries.csv",stringsAsFactors = FALSE)
regions<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/ETO Regions.csv",stringsAsFactors = FALSE)

bound<-function(x,l,h){
  data<-x
  data[x<l]<-l
  data[x>h]<-h
  return(data)
}

subset_plot<-function(data,size_group,track_group){
  ggplot(subset(data,data[,track_group]&fuel_size==size_group) %>%
    mutate(SEM=SEM*n/types,Custom=Custom*n/types,DI=DI*n/types,Standard=Standard*n/types,Lighting=Lighting*n/types) %>% 
    select(-n,-types) %>% 
    melt(id.vars = c("fuel_size","combo")))+
  geom_bar(aes(x=combo,y=value,fill=variable),stat = "identity",position="stack")+
  labs(x=NULL,y=NULL,fill=NULL,title=paste(size_group,track_group,sep = " "))+
  guides(fill=FALSE)+
  theme(axis.text.x=element_blank(),
    plot.title = element_text(size = 7),
    axis.text.y = element_text(size=6))
}

heat_data<-function(data=Proj_Pop,sector){
  data %>%
  filter(naicsgroup==sector) %>% 
  filter(participation=="Participant") %>% 
  group_by(fuel_size) %>% 
  summarise(
    SEM=mean(ifelse(SEM=="X",0,1)),
    Custom=mean(ifelse(Custom=="X",0,1)),
    DI=mean(ifelse(DI=="X",0,1)),
    Standard=mean(ifelse(Standard=="X",0,1)),
    Lighting=mean(ifelse(Lighting=="X",0,1))
    ) %>% 
  ungroup() %>% 
  melt(id.vars="fuel_size")
}

combo_data<-function(data=Proj_Pop,sector){
  data %>% 
  filter(naicsgroup==sector) %>% 
  filter(participation=="Participant") %>% 
  filter(combo!="") %>% 
  group_by(fuel_size,combo) %>% 
  summarise(n=n()) %>% 
  mutate(types=ifelse(combo=="",0,nchar(gsub("[^\r\n\t\f\v ]","",combo))+1),
    SEM=grepl("SEM",combo),
    Custom=grepl("Custom",combo),
    DI=grepl("DI",combo),
    Standard=grepl("Standard",combo),
    Lighting=grepl("Lighting",combo))
}

table_data<-function(data=Proj_Pop,sector,for_plot=FALSE){
  pop_dat<-data %>%
    filter(naicsgroup==sector) %>% 
    left_join(State_adj %>% filter(naicsgroup==sector),by="naicsgroup") %>% 
    summarise(Sites=unique(Units),
      GWh=(sum(kwh2017[participation=="Participant"],na.rm = TRUE)+sum(kwh2017[participation=="Non-Participant"]*adj[participation=="Non-Participant"],na.rm = TRUE))/1e6,
      M_Therms=(sum(therms2017[participation=="Participant"],na.rm = TRUE)+sum(therms2017[participation=="Non-Participant"]*adj[participation=="Non-Participant"],na.rm = TRUE))/1e6)
    
  part_dat<-data %>% 
    filter(naicsgroup==sector) %>% 
    filter(participation=="Participant") %>% 
    summarise(Sites=n_distinct(et_siteid),GWh=sum(kwh2017,na.rm = TRUE)/1e6,M_Therms=sum(therms2017,na.rm = TRUE)/1e6)
  
  nonpart_dat<-pop_dat-part_dat
  
  save_dat<-data %>%
    filter(naicsgroup==sector) %>% 
    summarise(Sites=sum(kWh_savings>0|Therms_savings>0,na.rm = TRUE),GWh=sum(kWh_savings,na.rm = TRUE)/1e6,M_Therms=sum(Therms_savings)/1e6)
  
  full_dat<-bind_rows(pop_dat,nonpart_dat,part_dat,save_dat)
  row.names(full_dat)<-c("Total (est)","Non-Participant", "Participant","Savings")
  
  if(for_plot==TRUE){
    full_dat$pop<-row.names(full_dat)
    plot_dat<-full_dat %>% filter(pop=="Participant"|pop=="Non-Participant") %>% melt(id.vars="pop")
    
    full_dat<-ggplot(plot_dat)+
      geom_bar(aes(x=variable,y=value,fill=pop),stat = "identity",position = "fill")+
      labs(y="Percentage",x="Metric",title="Distribution Between \nParticipants and Non-Participants",fill="Participation")+
      scale_fill_manual(
        values = c("#99CC33","darkorange"),
        breaks = c("Participant","Non-Participant"),
        labels = c("Participants","Non-Participants")
      )+
      scale_y_continuous(labels = scales::percent)+
      theme(
        text = element_text(size=8),
        axis.text = element_text(size=6),
        plot.title = element_text(size=10))+
      coord_flip()
  }
  
  return(full_dat)
}

map_plot<-function(data=Proj_Pop,sector){
  map_dat<-data %>% 
    filter(!grepl("multifamily",naicsgroup,ignore.case = TRUE)&naicsgroup!="Industrial") %>% 
    filter(participation=="Participant") %>% 
    group_by(Region) %>% 
    summarise(
      p_save=round(sum(kWh_savings[naicsgroup==sector])/sum(kWh_savings),3)*100,
      p_use=round(sum(as.numeric(kwh2017[naicsgroup==sector]),na.rm = TRUE)/sum(as.numeric(kwh2017),na.rm = TRUE),3)*100,
      save_ratio=ifelse(p_save<=.5&p_use<=.5,0,bound(log10(p_save/p_use),-2,2)),save_diff=bound(p_save-p_use,-10,10)
      )
  
  map<-map_data('county') %>% subset(region == 'oregon') %>% left_join(regions %>% mutate(merge=tolower(County)),by=c("subregion"="merge"))
  
  map_out<-left_join(map,map_dat,by=c("Regions.for.EB.Process"="Region"))
  
  output<-ggplot(map_out)+
    coord_quickmap(xlim = c(-124.6, -116.4), ylim = c(42, 46.3)) +
    geom_polygon(data = map_out, aes(x = long, y = lat, group = group,fill = save_diff), color = "black", size = .3)+
    # scale_fill_gradient2(mid="white",high="red",low="blue",midpoint = 0,limits=c(-3,3))+
    scale_fill_gradientn(colors = rev(c("#ff0000","#ff6666","white","#9ebbff","#004CFF")),limits=c(-10,10))+
    labs(fill="Savings % - \nUsage %",title="Regional Variation:\n Difference of Subsector Participant Savings and \nSubsector Participant Usage")+
    theme(axis.title = element_blank(),axis.text = element_blank(),axis.ticks = element_blank(),axis.line = element_blank(),
    text = element_text(size=8),
    plot.title = element_text(size=8))

  return(output)
}

comb_funct<-function(data=combos,track,metric="percent"){
  if(metric=="percent"){
    comb_out<-data[as.vector(data[,track]==TRUE),] %>% 
  group_by(Track=as.character(track),fuel_size) %>% 
  summarise(
    total=sum(n),
    SEM=ifelse(track=="SEM",sum(0,n[types==1])/total,sum(n*SEM)/total),
    Custom=ifelse(track=="Custom",sum(0,n[types==1])/total,sum(n*Custom)/total),
    DI=ifelse(track=="DI",sum(0,n[types==1])/total,sum(n*DI)/total),
    Standard=ifelse(track=="Standard",sum(0,n[types==1])/total,sum(n*Standard)/total),
    Lighting=ifelse(track=="Lighting",sum(0,n[types==1])/total,sum(n*Lighting)/total),
    Total=sum(n)/0
    ) %>% 
  select(-total) %>% 
  melt(id.vars=c("Track","fuel_size"))
    
    return(comb_out)
    }
  if(metric=="count"){
    comb_out<-data[as.vector(data[,track]==TRUE),] %>% 
  group_by(Track=as.character(track),fuel_size) %>% 
  summarise(
    total=sum(n),
    SEM=ifelse(track=="SEM",sum(0,n[types==1])/total,sum(n*SEM)/total),
    Custom=ifelse(track=="Custom",sum(0,n[types==1])/total,sum(n*Custom)/total),
    DI=ifelse(track=="DI",sum(0,n[types==1])/total,sum(n*DI)/total),
    Standard=ifelse(track=="Standard",sum(0,n[types==1])/total,sum(n*Standard)/total),
    Lighting=ifelse(track=="Lighting",sum(0,n[types==1])/total,sum(n*Lighting)/total),
    Total=sum(n)
    ) %>% 
  select(-total) %>% 
  melt(id.vars=c("Track","fuel_size")) %>% 
      mutate(value=ifelse(variable=="Total",round(value,0),paste(round(value*100,0),"%",sep="")))
    
  return(comb_out)
  }
}

pie_plot<-function(data=pie_dat){
  ggplot(data %>% filter(fuel_size!="Unknown"))+
  geom_bar(aes(x="",y=value,fill=variable),stat = "identity",position = "fill",width = 1,color="black",size=0)+
  # geom_text(aes(x="",y=value,label=round(value*100,0)))+
  coord_polar("y")+
  theme(
    text=element_text(size=14),
    legend.text = element_text(size=7),
    legend.title = element_text(size=8),
    legend.key.size = unit(.25,"inches"),
    legend.position = "left",
    axis.text = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_text(face = "bold",size=11),
    axis.ticks = element_blank(),
    strip.text.x = element_blank(),
    legend.margin=margin(0,0,0,0),
    legend.box.margin=margin(0,-4,0,0),
    plot.margin = margin(-10,0,-10,0),
    axis.line = element_blank(),
    plot.title = )+
  labs(fill="Tracks per \nSite",y="Number of Tracks per Site")+
  scale_fill_manual(
    values = c("#99CC33","darkorange","#999933","#666699","black"),
    breaks = c("N1","N2","N3","N4","N5"),
    labels = c("One","Two","Three","Four","Five"))+
  facet_grid(.~fuel_size)
}

comb_plot<-function(){
  ggplot(CombFull %>% filter(fuel_size!="Unknown") %>% 
    mutate(t_f=factor(Track,levels = c("SEM","Custom","DI","Standard","Lighting","Total"))))+
  geom_tile(aes(x=variable,fill=value*100,y=1))+
  geom_text(data=TextFull %>% filter(fuel_size!="Unknown") %>%
      mutate(t_f=factor(Track,levels = c("SEM","Custom","DI","Standard","Lighting","Total"))),
    aes(x=variable,y=1,label=value),size=3)+
  theme(text=element_text(size=10),
    axis.text.x = element_text(angle = 0, hjust = .5,size=6),
    axis.text.y = element_blank(),axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    strip.text.y = element_text(angle = 180),
    strip.text.x = element_text(size = 12))+
  scale_fill_gradient2(high = "red",low="green",mid = "yellow",midpoint = 50,na.value = "gray90",guide = FALSE)+
  scale_x_discrete(position = "top")+
  coord_fixed(ratio = 1)+
  labs(y=NULL,x=NULL,fill="% (by row)",title="Combinations of Tracks at Sites")+
  facet_grid(t_f~fuel_size,switch = "both")
}

freq_plot<-function(){
  ggplot(left_join(for_heat,size_count,by=c("fuel_size"="Var1")))+
  # geom_tile(aes(x=variable,y=fuel_size,fill=value*100),color="black")+
  geom_tile(aes(x=fuel_size,
    y=factor(variable,levels = rev(c("SEM","Custom","DI","Standard","Lighting","Total"))),
    fill=value*100),color="black")+
  scale_fill_gradient2(high = "#000000",low="#00ff00",mid="#008700",midpoint = 50,limit=c(0,100))+
  # geom_text(aes(x=variable,y=fuel_size,label=round(value*Freq,0)),size=2)+
  geom_text(aes(x=fuel_size,
    y=factor(variable,levels = rev(c("SEM","Custom","DI","Standard","Lighting","Total"))),
    label=round(value*Freq,0)),size=2)+
   labs(title="Project Type as a \nPercentage of Total Projects",y="Project Track",
    x="Site Size",fill="Percentage \n(of size group)")+
  theme(axis.text.x = element_text(angle = 0, hjust = .5,size = 5),
    text = element_text(size=6),
    axis.text.y=element_text(size=5),
    plot.title = element_text(size=6),
    plot.margin=grid::unit(c(0,0,0,0), "mm"))+
  coord_fixed(ratio=1)
}
```
#Title - ETO Subsector Summaries

##Some Summary Information Could Go Here

This might be some overall summary, or general trends, or explaination of how to read charts. 

### Cover Table
```{r cover_table}
CT<-Proj_Pop %>% group_by(Sector=naicsgroup) %>% summarise(Parts=n(),kWh=sum(kwh2017,na.rm = TRUE),Therms=sum(therms2017,na.rm = TRUE),
  kWh_Heat=round(sum(kWh_savings,na.rm = TRUE)/sum(Proj_Pop$kWh_savings,na.rm = TRUE)-sum(kwh2017,na.rm = TRUE)/sum(as.numeric(Proj_Pop$kwh2017),na.rm = TRUE),3)*100,
  Therms_Heat=round(sum(Therms_savings,na.rm = TRUE)/sum(Proj_Pop$Therms_savings,na.rm = TRUE)-sum(therms2017,na.rm = TRUE)/sum(Proj_Pop$therms2017,na.rm = TRUE),3)*100)


```


### Region Chart

### Combination Chart Key

##### page break

##Summary would probably have another page

This might be some additional summary, or general trends, or explaination of how to read charts. 

##### page break

```{r prep_plot_office,include=FALSE}

for_heat<-heat_data(sector = "Office")

size_count<-as.data.frame(table(Proj_Pop$fuel_size[Proj_Pop$naicsgroup=="Office"&Proj_Pop$participation=="Participant"]))

combos<-combo_data(sector = "Office")

CombSEM<-comb_funct(track = "SEM")
CombCust<-comb_funct(track = "Custom")
CombDI<-comb_funct(track = "DI")
CombStan<-comb_funct(track = "Standard")
CombLite<-comb_funct(track = "Lighting")

CombFull<-bind_rows(CombSEM,CombCust,CombDI,CombStan,CombLite)

TextSEM<-comb_funct(track = "SEM",metric = "count")
TextCust<-comb_funct(track = "Custom",metric = "count")
TextDI<-comb_funct(track = "DI",metric = "count")
TextStan<-comb_funct(track = "Standard",metric = "count")
TextLite<-comb_funct(track = "Lighting",metric = "count")

TextFull<-bind_rows(TextSEM,TextCust,TextDI,TextStan,TextLite)

pie_dat<-combos %>% 
  group_by(fuel_size) %>% 
  summarise(N1=sum(n[types==1])/sum(n),N2=sum(n[types==2])/sum(n),N3=sum(n[types==3])/sum(n),N4=sum(n[types==4])/sum(n),N5=sum(n[types==5])/sum(n)) %>% 
  melt(id.vars="fuel_size")

```

##Characterization of the Office Market

### Sector Overview

```{r off_tab,include=TRUE}
kable(table_data(sector="Office"),digits = 0)
```

```{r off_tab_plot,include=TRUE,fig.height=2,dpi=1000}
table_data(sector = "Office",for_plot = TRUE)

```




###Regional Variation in Savings Relative to Usage

Shows difference of regional subsector participant savings to regional subsector participant usage. Positive (red) values indicate disproportionately high savings, Negative (blue) values show disproportionately low savings.

```{r map_off,include=TRUE,fig.height=3,dpi=1000}
map_plot(sector = "Office")
```


###Plot of Track Frequency by Size

```{r track_freq_off,include=TRUE,fig.height=2,dpi=1000}
freq_plot()
```

###Plot of Track Combinations

```{r track_comb_off,include=TRUE,fig.width=7.5,dpi=2000,fig.height=3.25}
comb_plot()

```

```{r comb_count_off, include=TRUE,fig.width=6,dpi=2000,fig.height=2}
pie_plot(pie_dat)
```


```{r prep_plot_rest,include=FALSE}

for_heat<-heat_data(sector = "Restaurant")

size_count<-as.data.frame(table(Proj_Pop$fuel_size[Proj_Pop$naicsgroup=="Restaurant"&Proj_Pop$participation=="Participant"]))

combos<-combo_data(sector = "Restaurant")

CombSEM<-comb_funct(track = "SEM")
CombCust<-comb_funct(track = "Custom")
CombDI<-comb_funct(track = "DI")
CombStan<-comb_funct(track = "Standard")
CombLite<-comb_funct(track = "Lighting")

CombFull<-bind_rows(CombSEM,CombCust,CombDI,CombStan,CombLite)

TextSEM<-comb_funct(track = "SEM",metric = "count")
TextCust<-comb_funct(track = "Custom",metric = "count")
TextDI<-comb_funct(track = "DI",metric = "count")
TextStan<-comb_funct(track = "Standard",metric = "count")
TextLite<-comb_funct(track = "Lighting",metric = "count")

TextFull<-bind_rows(TextSEM,TextCust,TextDI,TextStan,TextLite)

pie_dat<-combos %>% 
  group_by(fuel_size) %>% 
  summarise(N1=sum(n[types==1])/sum(n),N2=sum(n[types==2])/sum(n),N3=sum(n[types==3])/sum(n),N4=sum(n[types==4])/sum(n),N5=sum(n[types==5])/sum(n)) %>% 
  melt(id.vars="fuel_size")
```
##### page break

##Characterization of the Restaurant Market

###Sector Overview

```{r rest_tab,include=TRUE}
kable(table_data(sector="Restaurant"),digits = 0)
```


```{r rest_tab_plot,include=TRUE, fig.height=2,dpi=1000}
table_data(sector="Restaurant",for_plot = TRUE)

```

###Regional Variation in Savings Relative to Usage

Shows difference of regional subsector participant savings to regional subsector participant usage. Positive (red) values indicate disproportionately high savings, Negative (blue) values show disproportionately low savings.

```{r map_rest,include=TRUE,fig.height=3,dpi=1000}
map_plot(sector = "Restaurant")
```

###Plot of Track Frequency by Size

```{r track_freq_rest,include=TRUE,fig.height=2,dpi=1000}
freq_plot()
```

###Plot of Track Combinations

```{r track_comb_rest,include=TRUE,fig.width=7.5,dpi=2000,fig.height=3.25}
comb_plot()
```

```{r comb_count_rest, include=TRUE,fig.width=6,dpi=2000,fig.height=2}
pie_plot(pie_dat)

```
