---
title: "ETO EB Sector Summaries"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```

```{r data, include=FALSE,warning=FALSE}
library(xlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(reshape2)
library(cowplot)
library(knitr)
library(maps)

Proj_Pop<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/Project_info_for_Summaries.csv",stringsAsFactors = FALSE)
State_adj<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/State_info_for_Summaries.csv",stringsAsFactors = FALSE)
regions<-read.csv("/volumes/Projects/430011 - ETO Existing Buildings/Data/ETO Regions.csv",stringsAsFactors = FALSE)

bound<-function(x,l,h){
  data<-x
  data[x<l]<-l
  data[x>h]<-h
  return(data)
  }

subset_plot<-function(data,size_group,track_group){
  ggplot(subset(data,data[,track_group]&fuel_size==size_group) %>%
    mutate(SEM=SEM*n/types,Custom=Custom*n/types,DI=DI*n/types,Standard=Standard*n/types,Lighting=Lighting*n/types) %>% 
    select(-n,-types) %>% 
    melt(id.vars = c("fuel_size","combo")))+
  geom_bar(aes(x=combo,y=value,fill=variable),stat = "identity",position="stack")+
  labs(x=NULL,y=NULL,fill=NULL,title=paste(size_group,track_group,sep = " "))+
  guides(fill=FALSE)+
  theme(axis.text.x=element_blank(),
    plot.title = element_text(size = 7),
    axis.text.y = element_text(size=6))
}

heat_data<-function(data=Proj_Pop,sector){
  data %>%
  filter(naicsgroup==sector) %>% 
  filter(participation=="Participant") %>% 
  group_by(fuel_size) %>% 
  summarise(
    SEM=mean(ifelse(SEM=="X",0,1)),
    Custom=mean(ifelse(Custom=="X",0,1)),
    DI=mean(ifelse(DI=="X",0,1)),
    Standard=mean(ifelse(Standard=="X",0,1)),
    Lighting=mean(ifelse(Lighting=="X",0,1))
    ) %>% 
  ungroup() %>% 
  melt(id.vars="fuel_size")
}

combo_data<-function(data=Proj_Pop,sector){
  data %>% 
  filter(naicsgroup==sector) %>% 
  filter(participation=="Participant") %>% 
  filter(combo!="") %>% 
  group_by(fuel_size,combo) %>% 
  summarise(n=n()) %>% 
  mutate(types=ifelse(combo=="",0,nchar(gsub("[^\r\n\t\f\v ]","",combo))+1),
    SEM=grepl("SEM",combo),
    Custom=grepl("Custom",combo),
    DI=grepl("DI",combo),
    Standard=grepl("Standard",combo),
    Lighting=grepl("Lighting",combo))
}

table_data<-function(data=Proj_Pop,sector){
  pop_dat<-data %>%
    filter(naicsgroup==sector) %>% 
    left_join(State_adj %>% filter(naicsgroup==sector),by="naicsgroup") %>% 
    summarise(Sites=unique(Units),
      GWh=(sum(kwh2017[participation=="Participant"],na.rm = TRUE)+sum(kwh2017[participation=="Non-Participant"]*adj[participation=="Non-Participant"],na.rm = TRUE))/1e6,
      M_Therms=(sum(therms2017[participation=="Participant"],na.rm = TRUE)+sum(therms2017[participation=="Non-Participant"]*adj[participation=="Non-Participant"],na.rm = TRUE))/1e6)
    
  part_dat<-data %>% 
    filter(naicsgroup==sector) %>% 
    filter(participation=="Participant") %>% 
    summarise(Sites=n_distinct(et_siteid),GWh=sum(kwh2017,na.rm = TRUE)/1e6,M_Therms=sum(therms2017,na.rm = TRUE)/1e6)
  
  save_dat<-data %>%
    filter(naicsgroup==sector) %>% 
    summarise(Sites=sum(kWh_savings>0|Therms_savings>0,na.rm = TRUE),GWh=sum(kWh_savings,na.rm = TRUE)/1e6,M_Therms=sum(Therms_savings)/1e6)
  
  full_dat<-bind_rows(pop_dat,part_dat,save_dat)
  row.names(full_dat)<-c("Total (est)","Participant","Savings")
  
  return(full_dat)
}

map_plot<-function(data=Proj_Pop,sector){
  map_dat<-data %>% 
    filter(!grepl("multifamily",naicsgroup,ignore.case = TRUE)&naicsgroup!="Industrial") %>% 
    filter(participation=="Participant") %>% 
    group_by(Region) %>% 
    summarise(
      p_save=round(sum(kWh_savings[naicsgroup==sector])/sum(kWh_savings),3)*100,
      p_use=round(sum(as.numeric(kwh2017[naicsgroup==sector]),na.rm = TRUE)/sum(as.numeric(kwh2017),na.rm = TRUE),3)*100,
      save_ratio=ifelse(p_save<=.5&p_use<=.5,0,bound(log10(p_save/p_use),-2,2)),save_diff=bound(p_save-p_use,-10,10)
      )
  
  map<-map_data('county') %>% subset(region == 'oregon') %>% left_join(regions %>% mutate(merge=tolower(County)),by=c("subregion"="merge"))
  
  map_out<-left_join(map,map_dat,by=c("Regions.for.EB.Process"="Region"))
  
  output<-ggplot(map_out)+
    coord_quickmap(xlim = c(-124.6, -116.4), ylim = c(42, 46.3)) +
    geom_polygon(data = map_out, aes(x = long, y = lat, group = group,fill = save_diff), color = "black", size = .3)+
    # scale_fill_gradient2(mid="white",high="red",low="blue",midpoint = 0,limits=c(-3,3))+
    scale_fill_gradientn(colors = rev(c("#ff0000","#ff6666","white","#9ebbff","#004CFF")),limits=c(-10,10))+
    labs(fill="Savings % - \nUsage %",title="Regional Variation:\n Difference of Subsector Participant Savings and \nSubsector Participant Usage")+
    theme(axis.title = element_blank(),axis.text = element_blank(),axis.ticks = element_blank(),axis.line = element_blank(),
    text = element_text(size=8),
    plot.title = element_text(size=8))

  return(output)
}

```

```{r prep_plot_office,include=FALSE}

for_heat<-heat_data(sector = "Office")

size_count<-as.data.frame(table(Proj_Pop$fuel_size[Proj_Pop$naicsgroup=="Office"&Proj_Pop$participation=="Participant"]))

combos<-combo_data(sector = "Office")

# ggplot(combos)+
#   geom_bar(aes(x=types,y=n,fill=combo),position = "fill",stat = "identity")+
#   facet_grid(fuel_size~.)
# 
# ggplot(combos)+
#   geom_bar(aes(x=fuel_size,y=n,fill=combo),position = "fill",stat = "identity")
# 
# ggplot(combos)+
#   geom_bar(aes(x=combo,y=n),stat="identity")+
#   facet_grid(fuel_size~.,scales = "free_y")+
#   theme(axis.text.x = element_text(angle = 90, hjust = .5))

combos2<-combos
combos2$SEM<-combos2$SEM*combos2$n/combos2$types
combos2$Custom<-combos2$Custom*combos2$n/combos2$types
combos2$DI<-combos2$DI*combos2$n/combos2$types
combos2$Standard<-combos2$Standard*combos2$n/combos2$types
combos2$Lighting<-combos2$Lighting*combos2$n/combos2$types

combos3<-melt(combos2 %>% select(-n,-types),id.vars = c("fuel_size","combo") )

LSEM<-subset_plot(combos,size_group = "Large",track_group = "SEM")
MSEM<-subset_plot(combos,size_group = "Medium",track_group = "SEM")
SSEM<-subset_plot(combos,size_group = "Small",track_group = "SEM")
LCUST<-subset_plot(combos,size_group = "Large",track_group = "Custom")
MCUST<-subset_plot(combos,size_group = "Medium",track_group = "Custom")
SCUST<-subset_plot(combos,size_group = "Small",track_group = "Custom")
LDI<-subset_plot(combos,size_group = "Large",track_group = "DI")
MDI<-subset_plot(combos,size_group = "Medium",track_group = "DI")
SDI<-subset_plot(combos,size_group = "Small",track_group = "DI")
LSTAN<-subset_plot(combos,size_group = "Large",track_group = "Standard")
MSTAN<-subset_plot(combos,size_group = "Medium",track_group = "Standard")
SSTAN<-subset_plot(combos,size_group = "Small",track_group = "Standard")
LLITE<-subset_plot(combos,size_group = "Large",track_group = "Lighting")
MLITE<-subset_plot(combos,size_group = "Medium",track_group = "Lighting")
SLITE<-subset_plot(combos,size_group = "Small",track_group = "Lighting")
USEM<-subset_plot(combos,size_group = "Unknown",track_group = "SEM")
UCUST<-subset_plot(combos,size_group = "Unknown",track_group = "Custom")
UDI<-subset_plot(combos,size_group = "Unknown",track_group = "DI")
USTAN<-subset_plot(combos,size_group = "Unknown",track_group = "Standard")
ULITE<-subset_plot(combos,size_group = "Unknown",track_group = "Lighting")
LEGEND<-ggplot(combos %>%
    mutate(SEM=SEM*n/types,Custom=Custom*n/types,DI=DI*n/types,Standard=Standard*n/types,Lighting=Lighting*n/types) %>% 
    select(-n,-types) %>% 
    melt(id.vars = c("fuel_size","combo")))+
  geom_bar(aes(x=combo,y=value,fill=variable),stat = "identity",position="stack")+
  labs(fill="Track")+
  theme(legend.text = element_text(size = 6),legend.title = element_text(size=8))

```

#Results for Offices

## Basic Table
```{r off_tab,include=TRUE}
kable(table_data(sector="Office"))
```

##Plot of Regional Variation in Savings
Shows difference of regional subsector participant savings to regional subsector participant usage. Positive (red) values indicate disproportionately high savings, Negative (blue) values show disproportionately low savings.

```{r map_off,include=TRUE,fig.height=3,dpi=1000}
map_plot(sector = "Office")
```


##Plot of Track Frequency by Size
```{r track_freq_off,include=TRUE,fig.height=3,dpi=1000}
ggplot(left_join(for_heat,size_count,by=c("fuel_size"="Var1")))+
  geom_tile(aes(x=variable,y=fuel_size,fill=value*100),color="black")+
  scale_fill_gradient2(high = "#000000",low="#00ff00",mid="#008700",midpoint = 50,limit=c(0,100))+
  geom_text(aes(x=variable,y=fuel_size,label=round(value*Freq,0)))+
   labs(title="Project Type as a \nPercentage of Total Projects",x="Project Track",
    y="Site Size",fill="Percentage \n(of size group)")+
  theme(axis.text.x = element_text(angle = 90, hjust = .5,size = 8),
    text = element_text(size=8),
    axis.text.y=element_text(size=8),
    plot.title = element_text(size=8))+
  coord_fixed(ratio=1)

```

##Plot of Track Combinations

```{r track_comb_off,include=TRUE,fig.height=5,dpi=1000}
plot_grid(
  plot_grid(
    LSEM,LCUST,LDI,LSTAN,LLITE,MSEM,MCUST,MDI,MSTAN,MLITE,SSEM,SCUST,SDI,SSTAN,SLITE,USEM,UCUST,UDI,USTAN,ULITE,
    ncol=5,nrow=4,label_size = 8
    # labels=c("Large SEM","Large Custom","Large DI","Large Standard","Large Lighting",
    #   "Medium SEM","Medium Custom","Medium DI","Medium Standard","Medium Lighting",
    #   "Small SEM","Small Custom","Small DI","Small Standard","Small Lighting",
    #   "Unknown SEM","Unknown Custom","Unknown DI","Unknown Standard","Unknown Lighting")
    ),
  plot_grid(get_legend(LEGEND),ncol=1),rel_widths = c(1,.15))


```

```{r prep_plot_rest,include=FALSE}

for_heat<-heat_data(sector = "Restaurant")

size_count<-as.data.frame(table(Proj_Pop$fuel_size[Proj_Pop$naicsgroup=="Restaurant"&Proj_Pop$participation=="Participant"]))

combos<-combo_data(sector = "Restaurant")

# ggplot(combos)+
#   geom_bar(aes(x=types,y=n,fill=combo),position = "fill",stat = "identity")+
#   facet_grid(fuel_size~.)
# 
# ggplot(combos)+
#   geom_bar(aes(x=fuel_size,y=n,fill=combo),position = "fill",stat = "identity")
# 
# ggplot(combos)+
#   geom_bar(aes(x=combo,y=n),stat="identity")+
#   facet_grid(fuel_size~.,scales = "free_y")+
#   theme(axis.text.x = element_text(angle = 90, hjust = .5))

combos2<-combos
combos2$SEM<-combos2$SEM*combos2$n/combos2$types
combos2$Custom<-combos2$Custom*combos2$n/combos2$types
combos2$DI<-combos2$DI*combos2$n/combos2$types
combos2$Standard<-combos2$Standard*combos2$n/combos2$types
combos2$Lighting<-combos2$Lighting*combos2$n/combos2$types

combos3<-melt(combos2 %>% select(-n,-types),id.vars = c("fuel_size","combo") )

LSEM<-subset_plot(combos,size_group = "Large",track_group = "SEM")
MSEM<-subset_plot(combos,size_group = "Medium",track_group = "SEM")
SSEM<-subset_plot(combos,size_group = "Small",track_group = "SEM")
LCUST<-subset_plot(combos,size_group = "Large",track_group = "Custom")
MCUST<-subset_plot(combos,size_group = "Medium",track_group = "Custom")
SCUST<-subset_plot(combos,size_group = "Small",track_group = "Custom")
LDI<-subset_plot(combos,size_group = "Large",track_group = "DI")
MDI<-subset_plot(combos,size_group = "Medium",track_group = "DI")
SDI<-subset_plot(combos,size_group = "Small",track_group = "DI")
LSTAN<-subset_plot(combos,size_group = "Large",track_group = "Standard")
MSTAN<-subset_plot(combos,size_group = "Medium",track_group = "Standard")
SSTAN<-subset_plot(combos,size_group = "Small",track_group = "Standard")
LLITE<-subset_plot(combos,size_group = "Large",track_group = "Lighting")
MLITE<-subset_plot(combos,size_group = "Medium",track_group = "Lighting")
SLITE<-subset_plot(combos,size_group = "Small",track_group = "Lighting")
USEM<-subset_plot(combos,size_group = "Unknown",track_group = "SEM")
UCUST<-subset_plot(combos,size_group = "Unknown",track_group = "Custom")
UDI<-subset_plot(combos,size_group = "Unknown",track_group = "DI")
USTAN<-subset_plot(combos,size_group = "Unknown",track_group = "Standard")
ULITE<-subset_plot(combos,size_group = "Unknown",track_group = "Lighting")
LEGEND<-ggplot(combos %>%
    mutate(SEM=SEM*n/types,Custom=Custom*n/types,DI=DI*n/types,Standard=Standard*n/types,Lighting=Lighting*n/types) %>% 
    select(-n,-types) %>% 
    melt(id.vars = c("fuel_size","combo")))+
  geom_bar(aes(x=combo,y=value,fill=variable),stat = "identity",position="stack")+
  labs(fill="Track")+
  theme(legend.text = element_text(size = 6),legend.title = element_text(size=8))

```

#Results for Restaurants

## Basic Table
```{r rest_tab,include=TRUE}
kable(table_data(sector="Restaurant"))
```

##Plot of Regional Variation in Savings
Shows log of the ratio of regional subsector participant savings to regional subsector participant usage. Positive (red) values indicate disproportionately high savings, Negative (blue) values show disproportionately low savings.

```{r map_rest,include=TRUE,fig.height=3,dpi=1000}
map_plot(sector = "Restaurant")
```

##Plot of Track Frequency by Size
```{r track_freq_rest,include=TRUE,fig.height=3,dpi=1000}
ggplot(left_join(for_heat,size_count,by=c("fuel_size"="Var1")))+
  geom_tile(aes(x=variable,y=fuel_size,fill=value*100),color="black")+
  scale_fill_gradient2(high = "#000000",low="#00ff00",mid="#008700",midpoint = 50,limit=c(0,100))+
  geom_text(aes(x=variable,y=fuel_size,label=round(value*Freq,0)))+
   labs(title="Project Type as a \nPercentage of Total Projects",x="Project Track",
    y="Site Size",fill="Percentage \n(of size group)")+
  theme(axis.text.x = element_text(angle = 90, hjust = .5,size = 8),
    text = element_text(size=8),
    axis.text.y=element_text(size=8),
    plot.title = element_text(size=8))+
  coord_fixed(ratio=1)

```

##Plot of Track Combinations

```{r track_comb_rest,include=TRUE,fig.height=5,dpi=1000}


# plot_grid(
#   plot_grid(
#     LSEM,LCUST,LDI,LSTAN,LLITE,MSEM,MCUST,MDI,MSTAN,MLITE,SSEM,SCUST,SDI,SSTAN,SLITE,USEM,UCUST,UDI,USTAN,ULITE,
#     ncol=5,nrow=4,label_size = 8
#     # labels=c("Large SEM","Large Custom","Large DI","Large Standard","Large Lighting",
#     #   "Medium SEM","Medium Custom","Medium DI","Medium Standard","Medium Lighting",
#     #   "Small SEM","Small Custom","Small DI","Small Standard","Small Lighting",
#     #   "Unknown SEM","Unknown Custom","Unknown DI","Unknown Standard","Unknown Lighting")
#     ),
#   plot_grid(get_legend(LEGEND),ncol=1),rel_widths = c(1,.15))


```
